Concurrent Programming in Harmony

### 2021-08-17 book/paper.tex
Consolidating appendices

\chapter{Value Types and Operators}
\label{ap:values}

@@

\section{Boolean}

@@

\section{Integer}

@@

\section{Atom}

@@

\section{Set}

@@

\section{Dictionary}

@@

\section{List or Tuple}

@@

\section{String}

@@

\section{Bag or Multiset}

@@

\section{Program Counter}

@@

\section{Address}

@@

\section{Context}

@@

\chapter{List of Statements}

@@

\section{Single expression evaluation}

@@

\section{Assignment}

@@

\section{\textbf{assert}}

@@

\section{\textbf{atomic}}

@@

\section{\textbf{await}}

@@

\section{\textbf{const}}

@@

\section{\textbf{def}}

@@

\section{\textbf{del}}

@@

\section{\textbf{for ... in [...where]}}

@@

\section{\textbf{from ... import}}

@@

\section{\textbf{go}}

@@

\section{\textbf{if ... [elif ...]* [else]}}

@@

\section{\textbf{import}}

@@

\section{\textbf{invariant}}

@@

\section{\textbf{let}}

@@

\section{\textbf{pass}}

@@

\section{\textbf{possibly}}

@@

\section{\textbf{select ... in}}

@@

\section{\textbf{sequential}}

@@

\section{\textbf{spawn}}

@@

\section{\textbf{trap}}

@@

\section{\textbf{while}}

@@

\chapter{List of Modules}
\label{ap:module}

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{Monitor()} & return a monitor mutex\\
\hline
\texttt{enter($m$)} & enter a monitor.  $m$ points to a monitor mutex\\
\hline
\texttt{exit($m$)} & exit a monitor.\\
\hline
\texttt{Condition()} & return a condition variable\\
\hline
\texttt{wait($c$, $m$)} & wait on condition variable pointed to by $c$ in monitor pointed to by $m$\\
\hline
\texttt{signal($c$, $m$)} & signal a condition variable\\
\hline
\end{tabular}

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{issubset($s$, $t$)} & returns whether $s$ is a subset of $t$ \\
\hline
\texttt{issuperset($s$, $t$)} & returns whether $s$ is a superset of $t$ \\
\hline
\texttt{add}($s$, $e$) & returns $s \cup \{ e \}$ \\
\hline
\texttt{remove}($s$, $e$) & returns $s - \{ e \}$ \\
\hline
\texttt{combinations}($b$, $k$) & returns set of all subsets of size $k$ \\
\hline
\end{tabular}

\vspace{1em}
For Python programmers: note that $s <= t$ does not check if $s$ is a subset of
$t$ when $s$ and $t$ are sets, as ``$<=$'' implements a total order on all Harmony values
including sets.

\chapter{List of Machine Instructions}
\label{ap:harmonybytecode}

\chapter{Contexts and Threads}
\label{app:context}

There is initially a single context with name
\texttt{\_\_init\_\_/()} and program counter 0.  It starts executing
in atomic mode until it finishes executing the last
\texttt{Return} instruction.
Other threads, created through \texttt{spawn} statements, do not
start executing until then.

\chapter{Harmony Language Details}
\label{ap:details}

The Harmony language borrows heavily from Python.  However, there are
some important differences that we will describe in this chapter.

@@

@@

The Harmony model checker, called \emph{Charm}, takes the output from
the compiler and explores the entire state space in breadth-first order.
Even Harmony does not really support input, there are three sources of
non-determinism that make this exploration non-trivial:

>>>>

\chapter{Harmony Language Details}

\section{Value Types and Operators}
\label{ap:values}

@@

\subsection*{Boolean}

@@

\subsection*{Integer}

@@

\subsection*{Atom}

@@

\subsection*{Set}

@@

\subsection*{Dictionary}

@@

\subsection*{List or Tuple}

@@

\subsection*{String}

@@

\subsection*{Bag or Multiset}

@@

\subsection*{Program Counter}

@@

\subsection*{Address}

@@

\subsection*{Context}

@@

\section{Statements}

@@

\subsection*{Single expression evaluation}

@@

\subsection*{Assignment}

@@

\subsection*{\textbf{assert}}

@@

\subsection*{\textbf{atomic}}

@@

\subsection*{\textbf{await}}

@@

\subsection*{\textbf{const}}

@@

\subsection*{\textbf{def}}

@@

\subsection*{\textbf{del}}

@@

\subsection*{\textbf{for ... in [...where]}}

@@

\subsection*{\textbf{from ... import}}

@@

\subsection*{\textbf{go}}

@@

\subsection*{\textbf{if ... [elif ...]* [else]}}

@@

\subsection*{\textbf{import}}

@@

\subsection*{\textbf{invariant}}

@@

\subsection*{\textbf{let}}

@@

\subsection*{\textbf{pass}}

@@

\subsection*{\textbf{possibly}}

@@

\subsection*{\textbf{select ... in}}

@@

\subsection*{\textbf{sequential}}

@@

\subsection*{\textbf{spawn}}

@@

\subsection*{\textbf{trap}}

@@

\subsection*{\textbf{while}}

@@

@@

\chapter{Modules}
\label{ap:module}

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{Monitor()} & return a monitor mutex\\
\hline
\texttt{enter($m$)} & enter a monitor.  $m$ points to a monitor mutex\\
\hline
\texttt{exit($m$)} & exit a monitor\\
\hline
\texttt{Condition()} & return a condition variable\\
\hline
\texttt{wait($c$, $m$)} & wait on condition variable pointed to by $c$ in monitor pointed to by $m$\\
\hline
\texttt{signal($c$, $m$)} & signal a condition variable\\
\hline
\end{tabular}

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{issubset($s$, $t$)} & returns whether $s$ is a subset of $t$ \\
\hline
\texttt{issuperset($s$, $t$)} & returns whether $s$ is a superset of $t$ \\
\hline
\texttt{add}($s$, $e$) & returns $s \cup \{ e \}$ \\
\hline
\texttt{remove}($s$, $e$) & returns $s \backslash \{ e \}$ \\
\hline
\texttt{combinations}($b$, $k$) & returns set of all subsets of size $k$ \\
\hline
\end{tabular}

\vspace{1em}
For Python programmers: note that $s <= t$ does not check if $s$ is a subset of
$t$ when $s$ and $t$ are sets, as ``$<=$'' implements a total order on all Harmony values including sets (and the subset relation is not a total order).

There is initially a single context with name
\texttt{\_\_init\_\_()} and program counter 0.  It starts executing
in atomic mode until it finishes executing the last
\texttt{Return} instruction.
Other threads, created through \texttt{spawn} statements, do not
start executing until then.

\section{Machine Instructions}
\label{ap:harmonybytecode}

\section{Contexts and Threads}
\label{app:context}

@@

The Harmony model checker, called \emph{Charm}, takes the output from
the compiler and explores the entire state space in breadth-first order.
Even though Harmony does not really support input, there are three sources of
non-determinism that make this exploration non-trivial:
